Tabular Constraint Learning

### 2016-04-07 Paper/ECAI/tabular.tex
Expanded pseudo code
new examples
formalization

\documentclass{ecai}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}

\usepackage[dvipsnames]{xcolor}
\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}

@@

\sergey{to himself we need structure here} 
\begin{figure}[thpb]
  \begin{center}
    \includegraphics[width=0.55\textwidth]{figures/Demo.png}
  \end{center}
  \vspace{-10pt}
  \caption{An example of constraint reconstruction (in blue) with indicated groups (in green)}
  \label{fig:main_example}
\end{figure}

@@

\newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
\newcommand{\generategroups}{\textit{generateGroups}\xspace}
\newcommand{\extractgroups}{\textit{extractGroups}\xspace}
\newcommand{\findassignment}{\textit{findAssignment}\xspace}

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $D$ -- dataset, (optional groups $G$)
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
    \If{$G$ is not defined}
       \State $G \gets \extractgroups(D)$
    \EndIf
  \For{$c \in \constraints$} \Comment{\constraints -- the set of predefined Excel constraints}
    \For{$G_1, \dots, G_n \in \generategroups(c,G)$}
      \State $S \gets S \cup \findassignment(c, G1, \dots, G_n)$
    \EndFor
  \EndFor\\
\Return $S$
\end{algorithmic}
\caption{First Order Model: Iterative Query Enumeration}
\label{algo:top_cycle}
\end{algorithm}

>>>>

\documentclass{ecai}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}

\usepackage[dvipsnames]{xcolor}
\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}
\newcommand{\samuel}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}

@@

\sergey{to himself we need structure here} 
\begin{figure}[thpb]
  \begin{center}
    \includegraphics[width=0.51\textwidth]{figures/Demo.png}
  \end{center}
  \vspace{-10pt}
  \caption{An example of constraint reconstruction (in blue) with indicated groups (in green)}
  \label{fig:main_example}
\end{figure}

@@

\newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
\newcommand{\format}[1]{\textit{#1}\xspace}
\newcommand{\generategroups}{\format{generateAssignments}}
\newcommand{\extractgroups}{\format{extractGroups}}
\newcommand{\extracttables}{\format{extractTables}}
\newcommand{\learnconstraints}{\format{learnConstraints}}
\newcommand{\findassignment}{\format{findSolutions}}
\newcommand{\postprocess}{\format{pruneRedundant}}

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $G$ -- set of groups
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
    \State $S \gets \emptyset$ \Comment{The set of solutions}
  \For{$c \in \constraints$} \Comment{\constraints -- the set of predefined Excel constraints}
  \State $v_1, ..., v_n =$ variables of $c$ 
    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S)$}
      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S)$
    \EndFor
  \EndFor\\
\Return $S$
\end{algorithmic}
\caption{Tabular constraint learning}
\label{algo:tcl}
\end{algorithm}

\sergey{we assume an order on the constraints (it doesn't have to be complete though)}

\subsection{Groups}
In this work we distinguish the following types of data: numeric and textual. The numeric type has two subtypes: integers and floats. We also consider the special element called \textit{None}, which has two types: numeric and textual. A set is called type-consistent iff all elements are either numeric or textual. Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.

A \textit{vector} is a subrange of a column or of a row that is type-consistent. If a vector is a subrange of a row (column), we say that it has a \textit{row} (\textit{column}) orientation. A \textit{group} is a subrange of vectors with the same orientation in a table. We use the following notation to refer to a row group $G$ in the $N$-th table with rows ranging from $a$ to $b$: $G = T_N[a{:}b,:]$, where $N,a,b$ are natural numbers. Similarly for a column group $G$ ranging from $a$ to $b$ columns in the $N$-th table we write $G = T_N[{:},a{:}b]$.

\subsection{Algorithm description}

Let us describe the key functions in Algorithm \ref{algo:tcl}.
\paragraph{$\generategroups(\textit{Constraint,GroupSet,Solutions})$} is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = SUM(X)} and $X,Y$ satisfy the following constraints \sergey{Samuel, can you add here, what is actually used for the sum now?}. Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint. However, many constraints have the same candidate generation procedures, e.g., min, max, avg, count, etc.

\paragraph{$\findassignment(\textit{Constraint,Candidates,Solutions})$} is the function looking for the subsets of vectors in the candidates satisfying the constraint. If multiple subsets satisfy the constraint, a maximal is selected. If $G_1,G_2$, associated with the variables $X,Y$ in \textit{Y = SUM(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and consecutive subset of vectors $x$ in $G_1$ such that the sum over $x$ is equal to $y$. For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}7]$ can be used for both $X$ and $Y$ in the row-wise sum constraint, then as $x$ would be selected $T_1[{:},3{:}6]$ and as $y$ would be $T_1[{:},7]$.

\paragraph{}

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $D$ -- dataset, (optional: tables $T$, groups $G$)
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
    \If{$T$ is \textbf{not} provided}
      \State $T \gets \extracttables(D)$
    \EndIf
    \If{$G$ is \textbf{not} provided}
      \State $G \gets \extractgroups(D, T)$
    \EndIf
    \State $S \gets \learnconstraints(G)$
    \State $S \gets \postprocess(S)$
    \State \Return $S$
\end{algorithmic}
\caption{Workflow}
\label{algo:workflow}
\end{algorithm}
